---
title: "part4"
output: pdf_document
date: "2023-11-21"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(dplyr)
library(tidyverse)
```

4. A simulation method to test other teams lineups against ours. This might be hard to actually do, and might require some simplification. For example, it might be tricky to choose every team's best set of athletes for them - so maybe we could find a way to compare the US team vs the best non-US team? And assign gold, silver, bronze in some weighted way based on if we beat them/if we lose, by how much? Probably need to get a little creative here. use an arbitrary lineup set for now and we can just run that on whatever our output from 2 is.

```{r}
# select gymnast team for other countries (assuming the gymnasts from other countries are randomly selected because it's impossible that all countries use identical method as US to select gymnasts to participate in the Olympic)
men_other_countries <- data.frame()
women_other_countries <- data.frame()
country_list <- unique(men_gym_data$Country)
country_list <- country_list[country_list != 'USA']
for (i in 1:length(country_list)) {
  data <- men_gym_data %>% 
    filter(Country == country_list[i])
  gymnasts <- unique(data$name)
  data = data.frame(country = country_list[i], gymnast = sample(gymnasts, min(5, length(gymnasts)))) # 5 gymnasts for each country
  men_other_countries <- bind_rows(men_other_countries, data)
}
for (i in 1:length(country_list)) {
  data <- women_gym_data %>% 
    filter(Country == country_list[i])
  gymnasts <- unique(data$name)
  data = data.frame(country = country_list[i], gymnast = sample(gymnasts, min(5, length(gymnasts)))) # 5 gymnasts for each country
  women_other_countries <- bind_rows(women_other_countries, data)
}
```

```{r}
# function that simulates a random score for a gymnast in an apparatus
simulate_score_men <- function(gymnast, apparatus) {
  data = men_gym_data %>% filter(name == gymnast, Apparatus == apparatus)
  if (nrow(data) == 0) { # if the gymnast has no historical data in this apparatus then return a score of 0
    return(0)
  }
  # else return simulated score
  data <- data%>% 
    group_by(name) %>% 
    summarise(mean = mean(Score, na.rm = T),
              var = var(Score, na.rm = T))
  mu <- data$mean
  var <- data$var
  score <- rnorm(1, mu, var)
  return(score)
}
simulate_score_women <- function(gymnast, apparatus) {
  data = women_gym_data %>% filter(name == gymnast, Apparatus == apparatus)
  if (nrow(data) == 0) { # if the gymnast has no historical data in this apparatus then return a score of 0
    return(0)
  }
  # else return simulated score
  data <- data%>% 
    group_by(name) %>% 
    summarise(mean = mean(Score, na.rm = T),
              var = var(Score, na.rm = T))
  mu <- data$mean
  var <- data$var
  score <- rnorm(1, mu, var)
  return(score)
}
```

```{r}
# individual competition results for each apparatus
apparatus_men <- unique(men_gym_data$Apparatus)
apparatus_women <- unique(women_gym_data$Apparatus)

num_iterations = 1000
total_medal_score = list() # store possible total medal scores for 1000 iterations
for (iter in 1:num_iterations) {
  total_medal = 0
  # men
  for (apparatus in apparatus_men) { # for each apparatus
    apparatus_medal = 0
    results <- data.frame() # competition results
    for (country in men_other_countries$country) { # each country can have 2 men competing in each final apparatus
      # other countries
      data = men_other_countries %>% filter(country == country)
      scores <- list()
      for (gymnast in data$gymnast) {
        score <- simulate_score_men(gymnast, apparatus)
        scores = append(scores, score)
      }
      if (length(scores) > 2) {
        scores <- lapply(scores, sort, decreasing=TRUE)
        scores <- scores[1:2]
      }
      data = data.frame(country = country, score = scores)
      results = bind_rows(results, data)
    }
    # us team
    us_scores <- list()
    for (gymnast in us_team_men) { # selected team members from part2
      score <- simulate_score_men(gymnast, apparatus)
      us_scores = append(us_scores, score)
    }
    if (length(us_scores) > 2) {
      us_scores <- lapply(us_scores, sort, decreasing=TRUE)
      us_scores <- us_scores[1:2]
    }
    data = data.frame(country = 'USA', score = us_scores)
    results = bind_rows(results, data)
    
    # sort scores for all countries
    results = results %>% 
      arrange(desc(score))
    # medals
    if (results[1,]$country == 'USA') {
      apparatus_medal = apparatus_medal + 3 # 3 for gold
    }
    else if (results[2,]$country == 'USA') {
      apparatus_medal = apparatus_medal + 2 # 2 for silver
    }
    else if (results[3,]$country == 'USA') {
      apparatus_medal = apparatus_medal + 1 # 1 for bronze
    }
  }
  total_medal = total_medal + apparatus_medal
  
  # women
  for (apparatus in apparatus_women) { # for each apparatus
    apparatus_medal = 0
    results <- data.frame() # competition results for each apparatus
    for (country in women_other_countries$country) { # each country can have 2 women competing in each final apparatus
      # other countries
      data = women_other_countries %>% filter(country == country)
      scores <- list()
      for (gymnast in data$gymnast) {
        score <- simulate_score_men(gymnast, apparatus)
        scores = append(scores, score)
      }
      if (length(scores) > 2) {
        scores <- lapply(scores, sort, decreasing=TRUE)
        scores <- scores[1:2]
      }
      data = data.frame(country = country, score = scores)
      results = bind_rows(results, data)
    }
    # us team
    us_scores <- list()
    for (gymnast in us_team_women) { # selected team members from part2
      score <- simulate_score_women(gymnast, apparatus)
      us_scores = append(us_scores, score)
    }
    if (length(us_scores) > 2) {
      us_scores <- lapply(us_scores, sort, decreasing=TRUE)
      us_scores <- us_scores[1:2]
    }
    data = data.frame(country = 'USA', score = us_scores)
    results = bind_rows(results, data)
    
    # sort scores for all countries
    results = results %>% 
      arrange(desc(score))
    # medals
    if (results[1,]$country == 'USA') {
      apparatus_medal = apparatus_medal + 3 # 3 for gold
    }
    else if (results[2,]$country == 'USA') {
      apparatus_medal = apparatus_medal + 2 # 2 for silver
    }
    else if (results[3,]$country == 'USA') {
      apparatus_medal = apparatus_medal + 1 # 1 for bronze
    }
  }
  total_medal = total_medal + apparatus_medal
  total_medal_score = c(total_medal_score, total_medal)
}
```

```{r}
# team all-around competition results

num_iterations = 1000
total_medal_score = list() # store total medal score for 1000 iterations
for (iter in 1:num_iterations) {
  total_medal = 0
  # men
  results <- data.frame() # competition results for all apparatus
  for (apparatus in apparatus_men) { # for each apparatus
     # competition results
    for (country in men_other_countries$country) { # each country 
      # other countries
      data = men_other_countries %>% filter(country == country)
      scores <- list()
      
      gymnasts <- data$gymnast
      if (length(gymnasts) > 4) {
        gymnasts <- sample(gymnasts, 4) # each country can select 4 gymnasts
      }
      for (gymnast in gymnasts) {
        score <- simulate_score_men(gymnast, apparatus)
        scores = append(scores, score)
      }
      if (legnth(scores) > 3) {
        scores <- lapply(scores, sort, decreasing=TRUE)
        scores <- scores[1:3] # 3 top scores out of 4 gymnasts
      }
      data = data.frame(country = country, score = sum(scores))
      results = bind_rows(results, data)
    }
    # us team
    us_scores <- list()
    for (gymnast in us_team_women) { # selected team members from part2
      score <- simulate_score_men(gymnast, apparatus)
      us_scores = append(us_scores, score)
    }
    gymnasts <- sample(gymnasts, 4) # each country can select 4 gymnasts
    for (gymnast in gymnasts) {
      score <- simulate_score_men(gymnast, apparatus)
      us_scores = append(us_scores, score)
    }
    us_scores <- lapply(us_scores, sort, decreasing=TRUE)
    us_score <- us_scores[1:3] # 3 top scores out of 4 gymnasts
    
    data = data.frame(country = 'USA', score = sum(us_scores))
    results = bind_rows(results, data)
  }
  
  results = results %>% 
    group_by(country) %>% 
    summarise(team_score = sum(score)) %>% # add up score for each apparatus
    arrange(desc(team_score))
  # medals
  if (results[1,]$country == 'USA') {
    total_medal = total_medal + 3 # 3 for gold
  }
  else if (results[2,]$country == 'USA') {
    total_medal = total_medal + 2 # 2 for silver
  }
  else if (results[3,]$country == 'USA') {
    total_medal = total_medal + 1 # 1 for bronze
  }
  
  # women
  results <- data.frame() # competition results for all apparatus
  for (apparatus in apparatus_men) { # for each apparatus
     # competition results
    for (country in men_other_countries$country) { # each country 
      # other countries
      data = men_other_countries %>% filter(country == country)
      scores <- list()
      
      gymnasts <- data$gymnast
      if (length(gymnasts) > 4) {
        gymnasts <- sample(gymnasts, 4) # each country can select 4 gymnasts
      }
      for (gymnast in gymnasts) {
        score <- simulate_score_men(gymnast, apparatus)
        scores = append(scores, score)
      }
      if (legnth(scores) > 3) {
        scores <- lapply(scores, sort, decreasing=TRUE)
        scores <- scores[1:3] # 3 top scores out of 4 gymnasts
      }
      data = data.frame(country = country, score = sum(scores))
      results = bind_rows(results, data)
    }
    # us team
    us_scores <- list()
    for (gymnast in us_team_women) { # selected team members from part2
      score <- simulate_score_men(gymnast, apparatus)
      us_scores = append(us_scores, score)
    }
    gymnasts <- sample(gymnasts, 4) # each country can select 4 gymnasts
    for (gymnast in gymnasts) {
      score <- simulate_score_men(gymnast, apparatus)
      us_scores = append(us_scores, score)
    }
    us_scores <- lapply(us_scores, sort, decreasing=TRUE)
    us_score <- us_scores[1:3] # 3 top scores out of 4 gymnasts
    
    data = data.frame(country = 'USA', score = sum(us_scores))
    results = bind_rows(results, data)
  }
  
  results = results %>% 
    group_by(country) %>% 
    summarise(team_score = sum(score)) %>% # add up score for each apparatus
    arrange(desc(team_score))
    # medals
  if (results[1,]$country == 'USA') {
    total_medal = total_medal + 3 # 3 for gold
  }
  else if (results[2,]$country == 'USA') {
    total_medal = total_medal + 2 # 2 for silver
  }
  else if (results[3,]$country == 'USA') {
    total_medal = total_medal + 1 # 1 for bronze
  }
  total_medal_score = c(total_medal_score, total_medal)
}
```

