---
title: "part4"
output: pdf_document
date: "2023-11-21"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(dplyr)
library(tidyverse)
```

4. A simulation method to test other teams lineups against ours. This might be hard to actually do, and might require some simplification. For example, it might be tricky to choose every team's best set of athletes for them - so maybe we could find a way to compare the US team vs the best non-US team? And assign gold, silver, bronze in some weighted way based on if we beat them/if we lose, by how much? Probably need to get a little creative here. use an arbitrary lineup set for now and we can just run that on whatever our output from 2 is.

```{r}
# function that simulates a random score for a gymnast in an apparatus
simulate_score_men <- function(gymnast, apparatus) {
  data = men_gym_data %>% filter(name == gymnast, Apparatus == apparatus)
  if (nrow(data) == 0) { # if the gymnast has no historical data in this apparatus then return a score of 0
    return(0)
  }
  # else return simulated score
  data <- data%>% 
    group_by(name) %>% 
    summarise(mean = mean(Score, na.rm = T),
              var = var(Score, na.rm = T))
  mu <- data$mean
  var <- data$var
  score <- rnorm(1, mu, var)
  return(score)
}
simulate_score_women <- function(gymnast, apparatus) {
  data = women_gym_data %>% filter(name == gymnast, Apparatus == apparatus)
  if (nrow(data) == 0) { # if the gymnast has no historical data in this apparatus then return a score of 0
    return(0)
  }
  # else return simulated score
  data <- data%>% 
    group_by(name) %>% 
    summarise(mean = mean(Score, na.rm = T),
              var = var(Score, na.rm = T))
  mu <- data$mean
  var <- data$var
  score <- rnorm(1, mu, var)
  return(score)
}
```


```{r selected us team}
# replace the righthand part with each us team selected from part 2 (here I use a random sample)
us_team_men <- sample(unique(us_men_gym_data$name), 5)
us_team_women <- sample(unique(us_womens_gym_data$name), 5)
```


```{r total medal score, warning=FALSE, message=FALSE}
# score3 for gold, 2 for silver, 1 for bronze
apparatus_men <- unique(men_gym_data$Apparatus)
apparatus_women <- unique(women_gym_data$Apparatus)
country_list <- unique(men_gym_data$Country)
country_list <- country_list[country_list != 'USA']
num_iterations = 100
total_medal_score = list() # store possible total medal scores for 1000 iterations
for (iter in 1:num_iterations) {
  total_medal = 0
  
  # select gymnast team for other countries (assuming the gymnasts from other countries are randomly selected because it's impossible that all countries use identical method as US to select gymnasts to participate in the Olympic)
  men_other_countries <- data.frame()
  women_other_countries <- data.frame()
  
  for (i in 1:length(country_list)) {
    data <- men_gym_data %>% 
      filter(Country == country_list[i])
    gymnasts <- unique(data$name)
    if (length(gymnasts) > 0) {
      data = data.frame(country = country_list[i], gymnast = sample(gymnasts, min(5, length(gymnasts)))) # at most 5 gymnasts for each country
      men_other_countries <- bind_rows(men_other_countries, data)
    }
    
  }
  for (i in 1:length(country_list)) {
    data <- women_gym_data %>% 
      filter(Country == country_list[i])
    gymnasts <- unique(data$name)
    if (length(gymnasts) > 0) {
      data = data.frame(country = country_list[i], gymnast = sample(gymnasts, min(5, length(gymnasts)))) # 5 gymnasts for each country
      women_other_countries <- bind_rows(women_other_countries, data)
    }
  }

  # men
  for (apparatus in apparatus_men) { # for each apparatus
    apparatus_medal = 0
    results <- data.frame() # competition results
    for (c in unique(men_other_countries$country)) { # each country can have 2 men competing in each final apparatus
      # other countries
      data = men_other_countries %>% filter(country == c)
      scores <- list()
      for (gymnast in data$gymnast) {
        score <- simulate_score_men(gymnast, apparatus)
        if (!is.na(score)) {
          scores = append(scores, score)
        }
      }
      scores <- scores[!is.na(scores)] # drop NA scores
      # assume gymnasts with highest simulated score are the 2 selected for individual finals
      if (length(scores) > 2) {
        scores <- lapply(scores, sort, decreasing=TRUE)
        scores <- scores[1:2]
      }
      scores = unlist(scores)
      if (length(scores) != 0) { # if there is at least 1 non-NA scores
        data = data.frame(country = c, score = scores)
        results = bind_rows(results, data)
      }
    }
    # us team
    us_scores <- list()
    for (gymnast in us_team_men) { # selected team members from part2
      score <- simulate_score_men(gymnast, apparatus)
      if (!is.na(score)) {
        us_scores = append(us_scores, score)
      }
    }
    us_scores <- us_scores[!is.na(us_scores)]
    if (length(us_scores) > 2) {
      us_scores <- lapply(us_scores, sort, decreasing=TRUE)
      us_scores <- us_scores[1:2]
    }
    us_scores = unlist(us_scores)
    if (length(us_scores) != 0) { # if there is at least 1 non-NA scores
      data = data.frame(country = 'USA', score = us_scores)
      results = bind_rows(results, data)
    }
    # sort scores for all countries
    results = results %>% 
      arrange(desc(score))
    # medals
    if (results[1,]$country == 'USA') {
      apparatus_medal = apparatus_medal + 3 # 3 for gold
    }
    else if (results[2,]$country == 'USA') {
      apparatus_medal = apparatus_medal + 2 # 2 for silver
    }
    else if (results[3,]$country == 'USA') {
      apparatus_medal = apparatus_medal + 1 # 1 for bronze
    }
  }
  total_medal = total_medal + apparatus_medal
  
  # women
  for (apparatus in apparatus_women) { # for each apparatus
    apparatus_medal = 0
    results <- data.frame() # competition results for each apparatus
    for (c in unique(women_other_countries$country)) { # each country can have 2 women competing in each final apparatus
      # other countries
      data = women_other_countries %>% filter(country == c)
      scores <- list()
      for (gymnast in data$gymnast) {
        score <- simulate_score_men(gymnast, apparatus)
        scores = append(scores, score)
      }
      if (length(scores) > 2) {
        scores <- lapply(scores, sort, decreasing=TRUE)
        scores <- scores[1:2]
      }
      scores <- unlist(scores)
      if (length(scores != 0)) {
        data = data.frame(country = c, score = scores)
        results = bind_rows(results, data)
      }
    }
    # us team
    us_scores <- list()
    for (gymnast in us_team_women) { # selected team members from part2
      score <- simulate_score_women(gymnast, apparatus)
      if (!is.na(score)) {
        us_scores = append(us_scores, score)
      }
    }
    if (length(us_scores) > 2) {
      us_scores <- lapply(us_scores, sort, decreasing=TRUE)
      us_scores <- us_scores[1:2]
    }
    us_scores <- unlist(us_scores)
    data = data.frame(country = 'USA', score = us_scores)
    results = bind_rows(results, data)
    
    # sort scores for all countries
    results = results %>% 
      arrange(desc(score))
    # medals
    if (results[1,]$country == 'USA') {
      apparatus_medal = apparatus_medal + 3 # 3 for gold
    }
    if (results[2,]$country == 'USA') {
      apparatus_medal = apparatus_medal + 2 # 2 for silver
    }
    if (results[3,]$country == 'USA') {
      apparatus_medal = apparatus_medal + 1 # 1 for bronze
    }
  }
  total_medal = total_medal + apparatus_medal
  
  # men team all-round
  results_team <- data.frame() # competition results for all apparatus
  for (apparatus in apparatus_men) { # for each apparatus
     # competition results
    for (country in unique(men_other_countries$country)) { # each country 
      # other countries
      data = men_other_countries %>% filter(country == country)
      scores <- list()
      
      gymnasts <- data$gymnast
      if (length(gymnasts) > 4) {
        gymnasts <- sample(gymnasts, 4) # each country can select 4 gymnasts
      }
      for (gymnast in gymnasts) {
        score <- simulate_score_men(gymnast, apparatus)
        scores = append(scores, score)
      }
      if (length(scores) > 3) {
        scores <- lapply(scores, sort, decreasing=TRUE)
        scores <- scores[1:3] # 3 top scores out of 4 gymnasts
      }
      scores = unlist(scores)
      if (length(scores) != 0) {
        data = data.frame(country = country, score = sum(scores))
        results_team = bind_rows(results_team, data)
      }
      
    }
    # us team
    us_scores <- list()
    for (gymnast in us_team_men) { # selected team members from part2
      score <- simulate_score_men(gymnast, apparatus)
      us_scores = append(us_scores, score)
    }
    gymnasts <- sample(gymnasts, 4) # each country can select 4 gymnasts
    for (gymnast in gymnasts) {
      score <- simulate_score_men(gymnast, apparatus)
      us_scores = append(us_scores, score)
    }
    us_scores <- lapply(us_scores, sort, decreasing=TRUE)
    us_scores <- unlist(us_scores[1:3]) # 3 top scores out of 4 gymnasts
    if (length(us_scores) != 0) {
      data = data.frame(country = 'USA', score = sum(us_scores))
      results_team = bind_rows(results_team, data)
    }
    
  }
  
  results_team = results_team %>% 
    group_by(country) %>% 
    summarise(team_score = sum(score)) %>% # add up score for each apparatus
    arrange(desc(team_score))
  # medals
  if (results_team[1,]$country == 'USA') {
    total_medal = total_medal + 3 # 3 for gold
  }
  if (results_team[2,]$country == 'USA') {
    total_medal = total_medal + 2 # 2 for silver
  }
  if (results_team[3,]$country == 'USA') {
    total_medal = total_medal + 1 # 1 for bronze
  }
  
  # women team all-round
  results_team <- data.frame() # competition results for all apparatus
  for (apparatus in apparatus_women) { # for each apparatus
     # competition results
    for (country in unique(women_other_countries$country)) { # each country 
      # other countries
      data = women_other_countries %>% filter(country == country)
      scores <- list()
      
      gymnasts <- data$gymnast
      if (length(gymnasts) > 4) {
        gymnasts <- sample(gymnasts, 4) # each country can select 4 gymnasts
      }
      for (gymnast in gymnasts) {
        score <- simulate_score_women(gymnast, apparatus)
        scores = append(scores, score)
      }
      if (length(scores) > 3) {
        scores <- lapply(scores, sort, decreasing=TRUE)
        scores <- scores[1:3] # 3 top scores out of 4 gymnasts
      }
      scores = unlist(scores)
      if (length(scores) != 0) {
        data = data.frame(country = country, score = sum(scores))
        results_team = bind_rows(results_team, data)
      }
      
    }
    # us team
    us_scores <- list()
    for (gymnast in us_team_women) { # selected team members from part2
      score <- simulate_score_women(gymnast, apparatus)
      us_scores = append(us_scores, score)
    }
    gymnasts <- sample(gymnasts, 4) # each country can select 4 gymnasts
    for (gymnast in gymnasts) {
      score <- simulate_score_women(gymnast, apparatus)
      us_scores = append(us_scores, score)
    }
    us_scores <- lapply(us_scores, sort, decreasing=TRUE)
    us_scores <- unlist(us_scores[1:3]) # 3 top scores out of 4 gymnasts
    if (length(us_scores) != 0) {
      data = data.frame(country = 'USA', score = sum(us_scores))
      results_team = bind_rows(results_team, data)
    }
    
  }
  
  results_team = results_team %>% 
    group_by(country) %>% 
    summarise(team_score = sum(score)) %>% # add up score for each apparatus
    arrange(desc(team_score))
    # medals
  if (results_team[1,]$country == 'USA') {
    total_medal = total_medal + 3 # 3 for gold
  }
  if (results_team[2,]$country == 'USA') {
    total_medal = total_medal + 2 # 2 for silver
  }
  if (results_team[3,]$country == 'USA') {
    total_medal = total_medal + 1 # 1 for bronze
  }
  
  # individual all-round
  results_indi <- data.frame() # competition results for all apparatus
  # men
  for (apparatus in apparatus_men) { # for each apparatus
     # competition results
    for (country in unique(men_other_countries$country)) { # each country 
      # other countries
      data = men_other_countries %>% filter(country == country)
      
      gymnasts <- data$gymnast
      if (length(gymnasts) > 2) {
        gymnasts <- sample(gymnasts, 2) # each country randomly select 2 gymnasts for individual all-round
      }
      for (gymnast in gymnasts) {
        score <- simulate_score_men(gymnast, apparatus)
        if (!is.na(score)) {
          data = data.frame(country = country, gymnast = gymnast, score = scores)
          results_indi = bind_rows(results_indi, data)
        }
      }
    }
    # us 
    gymnasts <- sample(us_team_men, 2) # randomly select 2 gymnasts from team selected in part2
    for (gymnast in gymnasts) {
      score <- simulate_score_men(gymnast, apparatus)
      data = data.frame(country = 'USA', gymnast = gymnast, score = us_scores)
      results_indi = bind_rows(results_indi, data)
    }
    
  }
  
  results_indi = results_indi %>% 
    group_by(country, gymnast) %>% 
    summarise(all_round_score = sum(score)) %>% # add up score for each apparatus
    arrange(desc(all_round_score))
  # medals
  if (results_indi[1,]$country == 'USA') {
    total_medal = total_medal + 3 # 3 for gold
  }
  if (results_indi[2,]$country == 'USA') {
    total_medal = total_medal + 2 # 2 for silver
  }
  if (results_indi[3,]$country == 'USA') {
    total_medal = total_medal + 1 # 1 for bronze
  }
  
  # women
  results_indi <- data.frame() # competition results for all apparatus
  for (apparatus in apparatus_women) { # for each apparatus
     # competition results
    for (country in unique(women_other_countries$country)) { # each country 
      # other countries
      data = women_other_countries %>% filter(country == country)
      
      gymnasts <- data$gymnast
      if (length(gymnasts) > 2) {
        gymnasts <- sample(gymnasts, 2) # each country randomly select 2 gymnasts for individual all-round
      }
      for (gymnast in gymnasts) {
        score <- simulate_score_women(gymnast, apparatus)
        if (!is.na(score)) {
          data = data.frame(country = country, gymnast = gymnast, score = scores)
          results_indi = bind_rows(results_indi, data)
        }
      }
    }
    # us 
    gymnasts <- sample(us_team_women, 2) # randomly select 2 gymnasts from team selected in part2
    for (gymnast in gymnasts) {
      score <- simulate_score_men(gymnast, apparatus)
      data = data.frame(country = 'USA', gymnast = gymnast, score = us_scores)
      results_indi = bind_rows(results_indi, data)
    }
    
  }
  
  results_indi = results_indi %>% 
    group_by(country, gymnast) %>% 
    summarise(all_round_score = sum(score)) %>% # add up score for each apparatus
    arrange(desc(all_round_score))
  # medals
  if (results_indi[1,]$country == 'USA') {
    total_medal = total_medal + 3 # 3 for gold
  }
  if (results_indi[2,]$country == 'USA') {
    total_medal = total_medal + 2 # 2 for silver
  }
  if (results_indi[3,]$country == 'USA') {
    total_medal = total_medal + 1 # 1 for bronze
  }
  total_medal_score = c(total_medal_score, total_medal)
}
```
```{r compare selected us teams}
# mean total medal score for the 100 simulations for one us team
mean(total_medal_score)
```

